### 2025-04-08 MetadataS1/MetadataS1.tex
add some comments of reviewer

Within terrestrial Arctic sites, the south plain of Bylot Island in the Canadian High Arctic (\textbf{Figure \ref{figure:food_web}}) hosts one of the longest and most intensive biodiversity monitoring programs \citep{gauthier2024a}. Monitoring on Bylot Island began in 1989 with a focus on the snow goose and it gradually expanded to other species over time. Currently, the program encompasses all significant vertebrate species in the community with continuous monitoring spanning more than a decade \citep{gauthier2024a}. Monitoring is also conducted at multiple spatial scales, including intensive and systematic observations conducted across a landscape spanning approximately 400 km\textsuperscript{2}. This approach allows the scaling of local density measurements to the landscape level when required and facilitates the estimation of abundance for less common and rare species.

@@

\subsection*{Objective}
Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that stay in the study area for a significant period of time, and excludes non-breeding individuals that stop for only a few days during their migration. We focus on adults, except for lemmings for which we have not distinguished between juveniles and adults. Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments. Additionally, we aim to provide the average body mass for each species in the community, enabling the conversion of abundances into biomasses.
\newpage
\section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
    \section*{A. Data set identity} Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut
    \section*{B. Data set identification codes}
     \texttt{BYLOT-species\_taxonomy.csv}\\
     \texttt{BYLOT-species\_abundance.csv}\\
     \texttt{BYLOT-species\_body\_mass.csv}\\
     \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

@@

    \section*{B. Specific subproject description}
        \subsection*{1. Site description}
                        \subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem in the eastern Canadian High-Arctic. An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:table_species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:table_species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.
                        \input{tables/table_species_name_strategy.tex}
%label-> table:table_species_name_strategy
\newpage
                \subsubsection*{b. Geography} Our 389 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
                \subsubsection*{c. Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
                \subsubsection*{d. Geology} See \citet{klassen1993} for a detailed description of the geology of the study area.
                \subsubsection*{e. Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Louis-Pierre Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
                \subsubsection*{f. Site history} See \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
                \subsubsection*{g. Climate} The mean annual air temperature since 1995 is -14.4°C, with mean seasonal temperature of 4.7°C in summer (June to August), -11°C in fall (September to November), -32.4°C in winter (December to February) and -19.4°C in spring (March to May; \citet{bylotclimate}). The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds \citep{gauthier2024a}. In summer, the study area received on average 77.5 mm of precipitation \citep{bylotclimate}. Additionnally, 102 days are frost-free annually on average  \citep{bylotclimate} and the study area typically remains free of snow from mid-June to late September \citep{gauthier2013}.
\newpage

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.6\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
\caption{Number of cackling goose nests found across the study area over time. The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016. The model presents a strong visual fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area.}
\label{figure:cackling}
\end{figure}

      \item[] \textit{\textbf{Snow goose}}\\
       Between 1999 and 2023, we assessed the abundance of snow geese in the study area through a multi-step process. We calculated the mean annual density of snow goose nests separately in the mesic and wetland habitats of the area occupied by the goose colony annually. We made slight adjustments to the goose colony perimeter defined from helicopter flights to include all snow goose point counts where at least one breeding pair had been observed (\textbf{Figure \ref{figure:sngo_goose_colony}}). To determine the mean density of nesting geese in wetlands, we divided two times (assuming two individuals per nest) the total number of nests found during systematic nest searches by the total area of wetlands sampled. The density of geese nesting in mesic habitat, a less preferred nesting habitat \citep{lecomte2008}, was averaged from three independent methods: systematic nest searches, vertebrate count transects, and snow goose point counts. Systematic nest searches were highly precise, but covered a relatively small area, whereas transects and snow goose point counts were less precise but covered larger areas. For each method, we calculated the mean density of breeding individuals in mesic habitat by dividing the number of birds (or nests) recorded by the area sampled. Despite methodological differences, the three approaches showed similar inter-annual variations, supporting the use of a mean values to estimate nest density in mesic habitat \textbf{Figure \ref{figure:sngo_interannual_variation}}). Lastly, to transform the densities in total abundance, we determined the annual proportion of wetland and mesic habitats within the goose colony and multiplied the area of each habitat by the density of breeding individuals. For the period 1999 to 2009, we used the average limits of the colony over the period 2010 to 2023 because we did not conduct aerial survey of the colony. Moreover, nest density in the mesic habitat was derived from a single method (\textbf{Figure \ref{figure:sngo_interannual_variation}}).

@@

                \item[] \textit{\textbf{Rock ptarmigan}}\\
                We estimated the abundance of rock ptarmigans based on the annual nest density measured in the 8 km\textsuperscript{2} extensive nest search area of the Qarlikturvik valley. While we acknowledge that the opportunistic monitoring of this species likely underestimates nest density, the extensive sampling effort deployed annually within this area gives us confidence in obtaining a realistic number of nests. We then extrapolate the density to the whole study area, without distinction between mesic, wetland and upland habitats (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat. To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest).

@@

                \item[] \textit{\textbf{Arctic fox}}\\
                We estimated the abundance of Arctic foxes in the study area based on their estimated home range size inside and outside the goose colony. We used the data and methodologies outlined in \citet{dulude2023} to estimate home range size. However, here, we did not account for annual variations in lemming density as presented in \citet{dulude2023} in order to obtain mean fox home range. Given that foxes are territorial and exhibit an average spatial overlap with adjacent territories of 18\% \citep{clermont2021}, we converted home range size into individual density using the following formula: $\text{\textit{density of individuals}}= \frac{2}{0.82\times \text{\textit{home range}}}$. We used two as numerator because we assumed each territory was held by a pair of fox, either breeding or non-breeding fox pair, without accounting for potential nomadic or transient individuals. We used values of 12.26 km\textsuperscript{2} to represent the mean home range of foxes within the goose colony and 20.02 km\textsuperscript{2} for foxes outside the goose colony. We estimated the mean density of foxes in each zone of the study area according to the mean proportion of the zone covered by the goose colony. We derived the mean annual proportion of each zone covered by the goose colony from the colony outline between 2010 and 2023. We estimated a mean density of 0.14 individuals/km\textsuperscript{2} for the study area. Previously, the minimum density of foxes in the study area was estimated to be between 0.03 and 0.13 individuals per km\textsuperscript{2} based on camera traps \citep{royerboutin2015}. 
                \newpage
                        \end{enumerate}

@@

   \section*{Acknowledgements}
We are deeply appreciative of the extensive data gathered over the decades by generations students and researchers during their fieldwork at the Bylot Island research station, which made this project possible. Our gratitude also extends to the Center of Northern Studies for providing research facilities, as well as to the Polar Continental Shelf Program and Sirmilik National Park (Parks Canada) for their logistical support throughout the years. Additionally, we would like to express our special thanks to the Mittimatalik community and the Mittimatalik Hunters and Trappers Organization for their ongoing support of ecological monitoring on Bylot Island and for permitting us to conduct research on their land.

>>>>

The south plain of Bylot Island in the Canadian High Arctic (\textbf{Figure \ref{figure:food_web}}) hosts one of the longest and most intensive vertebrate monitoring programs in the High-Arctic \citep{gauthier2024a}. Monitoring on Bylot Island began in 1989 with a focus on the snow goose and it gradually expanded to other species over time. The program currently encompasses all vertebrate species in the community that have a significant influence on the food web dynamics \citep{gauthier2011, legagneux2012}, with continuous monitoring spanning more than a decade \citep{gauthier2024a}. Monitoring is also conducted at multiple spatial scales, including intensive and systematic observations conducted across a landscape spanning approximately 400 km\textsuperscript{2}. This approach allows the scaling of local density measurements to the landscape level when required and facilitates the estimation of abundance for less common and rare species.

@@

\subsection*{Objective}
Our main objective is to provide readily accessible, long-term time series of annual abundances of all vertebrate species within the Arctic terrestrial community of Bylot Island during the breeding season (May to August). This includes both breeding and non-breeding individuals that remain in the study area for a significant period of time (e.g., territorial breeding or non-breeding foxes), and excludes non-breeding individuals that stop only briefly during migration (e.g., shorebirds using Bylot Island for a short stopover). We focus on the breeding season here due to the absence of monitoring during the non-breeding season; however, we note that ecological dynamics during the non-breeding period, for both resident \citep{hutchison2020} and migratory species \citep{moisan2023}, can influence the food web during the breeding season. We focus on adults, except for lemmings for which we have not distinguished between juveniles and adults. Our focus extends to estimating abundances at the landscape scale, enabling the study of community and ecosystem dynamics, trophic interactions and the impacts of global changes on high-latitude environments. Additionally, we aim to provide the average body mass for each species in the community, enabling the conversion of abundances into biomasses.
\newpage
\section*{\textcolor{Blue}{Class I. Data Set Descriptors}}
    \section*{A. Data set identity} Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut
    \section*{B. Data set identification codes}
     \texttt{BYLOT-species\_taxonomy.csv}\\
     \texttt{BYLOT-species\_abundance.csv}\\
     \texttt{BYLOT-species\_body\_mass.csv}\\
     \texttt{BYLOT-interannual\_variation\_nest\_density.csv}\\

@@

    \section*{B. Specific subproject description}
        \subsection*{1. Site description}
                        \subsubsection*{a. Site type}
  The study area (389 km\textsuperscript{2}) represents a relatively productive tundra ecosystem compared to other sites at similar latitudes in the eastern Canadian High Arctic \citep{gauthier2024a}. An important biological characteristic of the area is the presence of a large snow goose (scientific names of most vertebrate species can be found in \textbf{Table \ref{table:table_species_name_strategy}}) colony of around 25 000 breeding pairs \citep{reed2002} spanning approximately 70 km\textsuperscript{2}. The vertebrate community within the study area comprises 30 bird species, with 29 of them being migratory or partially migratory, along with 5 mammal species (\textbf{Table \ref{table:table_species_name_strategy}}; \citet{moisan2023, gauthier2024a}). The study area experiences significant temporal fluctuations in the population of small mammals (lemmings), which in turn impact the occurrence and abundance of their avian and mammalian predators such as snowy owls, rough-legged hawks, long-tailed jaegers, ermines and Arctic foxes \citep{legagneux2012, duchesne2021}. We exclude occasional visitors, namely: i) species lacking confirmed breeding occurrences on the study site, ii) species observed solely within a single year, and iii) species primarily breeding and foraging in nearby marine or coastal habitats \citep{moisan2023}. The case of the red fox (\textit{Vulpes vulpes}) was ambiguous. While the presence of breeding pairs has been confirmed in the study area \citep{lai2022}, the extent of population establishment remains unclear and sightings are rare. Therefore, we decided to exclude this species.
                        \input{tables/table_species_name_strategy.tex}
%label-> table:table_species_name_strategy
                \subsubsection*{b. Geography} Our 389 km\textsuperscript{2} study area is located on the southern plain of Bylot Island, Nunavut, Canada (72.889 N, -79.906 W; \textbf{Figure \ref{figure:zones}}).
                \subsubsection*{c. Habitat}  The study area comprises a combination of mesic tundra mainly on hills (64 \%), upland plateaus of sedimentary rock with drier/rockier habitat at higher elevation (20 \%), low-lying wetlands interspersed with ponds (10 \%) and larger bodies of water such as lakes and rivers (6 \%).
                \subsubsection*{d. Geology} See \citet{klassen1993} for a detailed description of the geology of the study area.
                \subsubsection*{e. Hydrology} Wetlands were delineated by photo-interpretation of high-resolution satellite images (30 cm; Louis-Pierre Ouellet, unpublished data), whereas lakes were delineated with aerial photos and rivers with google satellite images, resulting in a coarser delineation.
                \subsubsection*{f. Site history} See \citet{gauthier2024a, gauthier2024b}  for a complete and detailed history of the site.
                \subsubsection*{g. Climate} The mean annual air temperature since 1995 is -14.4°C, with mean seasonal temperature of 4.7°C in summer (June to August), -11°C in fall (September to November), -32.4°C in winter (December to February) and -19.4°C in spring (March to May; \citet{bylotclimate}). The climate of the southern plain of Bylot Island is generally milder than that of the surrounding latitudes, as the plain present a southern exposure and the mountains to the north protect the plain from cold northerly winds \citep{gauthier2024a}. In summer, the study area received on average 77.5 mm of precipitation. On average, snow depth is relatively thin in December, January, and February ($\sim$15 cm) and increases from March, reaching its peak in late May ($\sim$27 cm) before decreasing until mid-June ($\sim$3 cm) \citep{bylotclimate}. The study area typically remains free of snow from mid-June to late September \citep{gauthier2013} with an average of 102 frost-free days annually \citep{bylotclimate}. From September to December, snow depth typically increases continuously \citep{bylotclimate}. The summary statistics presented here on the climate of the South Plain of Bylot Island are derived from automatic weather stations located in the Qarlikturvik Valley, which have been recording data on air temperature, thawing degree days, frost-free days, summer precipitation, snow thickness, and wind speed since 1994 \citep{bylotclimate}. Raw data from these weather stations are publicly accessible at \url{https://nordicana.cen.ulaval.ca/fr/publication.php?doi=45039SL-EE76C1BDAADC4890}.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.6\linewidth]{figures/cackling_goose_nest_exponential.pdf} 
\caption{Number of cackling goose nests found across the study area over time. The red line represents the fitted model over the period 1996-2023 ($y= e^{0.1717x-342.684}$), which was used to estimate the annual abundance of cackling goose between 1996 and 2016. The model presents a strong visual fit with the number of nests found between 2017 and 2023, when nest monitoring was systematic across the study area.}
\label{figure:cackling}
\end{figure}
      \item[] \textit{\textbf{Snow goose}}\\
       Between 1999 and 2023, we assessed the abundance of snow geese in the study area through a multi-step process. We calculated the mean annual density of snow goose nests separately in the mesic and wetland habitats of the area occupied by the goose colony annually. We made slight adjustments to the goose colony perimeter defined from helicopter flights to include all snow goose point counts where at least one breeding pair had been observed (\textbf{Figure \ref{figure:sngo_goose_colony}}). To determine the mean density of nesting geese in wetlands, we divided two times (assuming two individuals per nest) the total number of nests found during systematic nest searches by the total area of wetlands sampled. The density of geese nesting in mesic habitat, a less preferred nesting habitat \citep{lecomte2008}, was averaged from three independent methods: systematic nest searches, vertebrate count transects, and snow goose point counts \textbf{Figure \ref{figure:sngo_interannual_variation}}). Systematic nest searches were highly precise, but covered a relatively small area, whereas transects and snow goose point counts were less precise but covered larger areas. For each method, we calculated the mean density of breeding individuals in mesic habitat by dividing the number of birds (or nests) recorded by the area sampled. We found a strong and statistically significant correlation between density estimates derived from vertebrate count transects and snow goose point counts (Sperman's rank correlation r\textsubscript{s}= 0.8; p=0.003). However, there was no significant correlation between density estimates from nest sampling plots and either the transect data (r\textsubscript{s}=0.37; p=0.24) or the point counts (r\textsubscript{s}=0.22; p= 0.5). Despite the lack of correlation between observations (transects and point counts) and nest sampling, we chose to average the estimates from all three methods since observations and nest monitoring present complementary strengths and limitations. Transects and point counts offer broader spatial coverage but lower precision, while nest sampling provides higher precision but limited spatial coverage. Averaging the three methods allows us to balance these trade-offs and produce estimates that are less precise individually, but likely more accurate overall. Lastly, to transform the densities in total abundance, we determined the annual proportion of wetland and mesic habitats within the goose colony and multiplied the area of each habitat by the density of breeding individuals. For the period 1999 to 2009, we used the average limits of the colony over the period 2010 to 2023 because we did not conduct aerial survey of the colony. Moreover, nest density in the mesic habitat was derived from a single method (\textbf{Figure \ref{figure:sngo_interannual_variation}}).

@@

                \item[] \textit{\textbf{Rock ptarmigan}}\\
                We estimated the abundance of rock ptarmigans based on the annual nest density measured in the 8 km\textsuperscript{2} extensive nest search area of the Qarlikturvik valley. While we acknowledge that the opportunistic monitoring of this species likely underestimates nest density, the extensive sampling effort deployed annually within this area gives us confidence in obtaining a realistic number of nests. We then extrapolate the density to the whole study area, without distinction between mesic, wetland and upland habitats (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Among the 6 nests found in the study area, 4 were located in mesic habitat, while one nest was found in a wetland and another in an upland habitat. Nest site selection by rock ptarmigan has not been well documented in the Canadian Arctic; however, a study conducted at Sarcpa Lake (n= 7 nests) also highlight that rock ptarmigan can nests in a variety of habitats \citep{montgomerie1983}. To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest).

@@

                \item[] \textit{\textbf{Arctic fox}}\\
                We estimated the abundance of Arctic foxes in the study area based on their estimated home range size inside and outside the goose colony. We used the data and methodologies outlined in \citet{dulude2023} to estimate home range size of territorial foxes. However, here, we did not account for annual variations in lemming density as presented in \citet{dulude2023} in order to obtain mean fox home range. Given that foxes are territorial and exhibit an average spatial overlap with adjacent territories of 18\% \citep{clermont2021}, we converted home range size into individual density using the following formula: $\text{\textit{density of individuals}}= \frac{2}{0.82\times \text{\textit{home range}}}$. We used two as numerator because we assumed each territory was held by a pair of fox, either breeding or non-breeding fox pair, without accounting for potential nomadic or transient individuals. We used values of 12.26 km\textsuperscript{2} to represent the mean home range of foxes within the goose colony and 20.02 km\textsuperscript{2} for foxes outside the goose colony. We estimated the mean density of foxes in each zone of the study area according to the mean proportion of the zone covered by the goose colony. We derived the mean annual proportion of each zone covered by the goose colony from the colony outline between 2010 and 2023. We estimated a mean density of 0.14 individuals/km\textsuperscript{2} for the study area. Previously, the minimum density of foxes in the study area was estimated to be between 0.03 and 0.13 individuals per km\textsuperscript{2} based on camera traps \citep{royerboutin2015}. 
                \newpage
                        \end{enumerate}

@@

   \section*{Acknowledgements}
We are deeply appreciative of the extensive data gathered over the decades by generations of students and researchers during their fieldwork at the Bylot Island research station, which made this project possible. Our gratitude also extends to the Center of Northern Studies for providing research facilities, as well as to the Polar Continental Shelf Program and Sirmilik National Park (Parks Canada) for their logistical support throughout the years. Additionally, we would like to express our special thanks to the Mittimatalik community and the Mittimatalik Hunters and Trappers Organization for their ongoing support of ecological monitoring on Bylot Island and for permitting us to conduct research on their land.
